---
title: "Exploratory data analysis of House Prices dataset"
author: "Alexandra Kondekova"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(miscset)

# data
houses <- read.csv('train.csv')
```
First look at the data:
```{r}
str(houses)
dim(houses)
head(houses)
houses$MSSubClass <- as.factor(houses$MSSubClass)
```

Let us look at the missing data:
```{r}
miss <- apply(houses, 2, is.na) %>% apply(., 2, sum)
miss[miss > 0.5 * nrow(houses)]  # more than 50% of data is NA
```
However, in all cases, where NAs are more than 50% of the data, they do not mean missing values. They represent the 'no'/'none' categories in the variables. In the analysis, consider merging the categories to create binary variables (yes and no).

The analysis of continuous (numerical) variables
```{r}
num.vars <- Filter(is.numeric, houses) %>% names()  # 37
num.vars <- num.vars[-1]  # delete id

summary(houses[num.vars])
```
We see that some variables have only few possible values (e.g. BsmtFullBath, BsmtHalfBath, FullBath, HalfBath,KitchenAbvGr, Fireplaces, MoSoldby). In further analysis, these variables may be considered as factors.

Let us now look at the correlation structure of numerical variables.

```{r, fig.align = "center"}
corr <- cor(houses[num.vars]) %>% round(2)
ggcorrplot(corr, type = 'lower')
```

We see high correlation between variables that are obviously linked together, e.g. GarageCars and GarageArea, YearBuild and YearRemodAd, etc.

The analysis of categorical variables

```{r}
fac.vars <- Filter(is.factor, houses) %>% names ()  #44

summary(houses[fac.vars])
```

```{r, echo = FALSE, fig.align = "center", fig.height=15, fig.width=15}
ggplotGrid(ncol = 4,
           lapply(fac.vars[1:22],
                  function(col) {
                    ggplot(houses, aes_string(col)) + geom_bar() + coord_flip()
                  }))
ggplotGrid(ncol = 4,
           lapply(fac.vars[23:44],
                  function(col) {
                    ggplot(houses, aes_string(col)) + geom_bar() + coord_flip()
                  }))
```

Variable Utilities is totally useless as it contains only two categories and in one of them only one observation. Condition2 might be also considered as useless - small numbers of observations in some categories. In further analysis, consider merging categories for some variables as some of them contain only few observations (e.q. quality).

Let us look more closely at the target variable SalePrices.

```{r}
summary(houses$SalePrice)
```

This is the histogram and the estimation fo its density:
```{r, echo = FALSE, fig.align = "center"}
ggplot(houses, aes(x = SalePrice)) + 
  geom_histogram(aes(y = ..density..), colour = "black", fill = "#E69F00") +
  geom_density(alpha=.2, fill = "#999999")
```

We see that the distribution is skewed. 

Now, we plot the boxplots for sale prices per categories of all categorical variables.

```{r, echo = FALSE, fig.align = "center", fig.height=15, fig.width=15}
ggplotGrid(ncol = 4,
           lapply(fac.vars[1:22],
                  function(col) {
                    ggplot(houses, aes_string(x = col, y = "SalePrice")) + geom_boxplot()
                  }))

ggplotGrid(ncol = 4,
           lapply(fac.vars[23:44],
                  function(col) {
                    ggplot(houses, aes_string(x = col, y = "SalePrice")) + geom_boxplot()
                  }))
```

Now, we look at the correlation of sale prices with the rest of numerical variables.

```{r}
(cor.sp <- corr[, "SalePrice"])
```
```{r, echo = FALSE}
cor.sp.high <- cor.sp[abs(cor.sp) >= 0.5 & !is.na(cor.sp)]
```

These are the variables where the correlation with sale prices is higher than 0.5. 
```{r}
cor.sp.high
```
```{r, echo = FALSE}
cor.sp.high <- cor.sp.high[-length(cor.sp.high)] %>% names
```

Let us plot the highly correlated variables versus the sale prices.

```{r, echo = FALSE, fig.align = "center", fig.height=15, fig.width=15}
ggplotGrid(ncol = 2,
           lapply(cor.sp.high,
                  function(col) {
                    ggplot(houses, aes_string(x = col, y = "SalePrice")) +
                    geom_point()
                  }))
```

We see some trends.

